A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker

### 2020-03-15 manuscript.Rmd
mention GitLab closes #13

Git also makes it particularly easy to share and collaborate on a project with other researchers. A popular service for sharing materials via Git is [GitHub](https://github.com). Just sharing Git repositories with the public is always free, private repositories (only visible to persons you invite) are [free for researchers](https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount). After creating a user account, one can create a new repository and GitHub provides information on how to upload your repository from the terminal, e.g., for our repository (here with user name 'aaronpeikert' and repository name 'reproducible-research'):

@@

While our approach was designed to scale well with the complexity of a computing-intense project, we realise that this flexibility may not be straightforward to integrate into a researcher's everyday workflow. There are R packages that implement parts of our workflow and, thus, lower the threshold for adoption when the full flexibility provided by our workflow is not needed. The use of R Markdown within a project, tracked with Git can be simplified with the [`workflowR`](https://github.com/jdblischak/workflowr) package [@workflowr]. The [`drake`](https://github.com/ropensci/drake) package [@drake] is directly inspired by Make and takes an R centric approach, making it especially suited for projects only involving R, but can also handle external dependencies. The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [`holepunch`](https://karthik.github.io/holepunch/) package [@holepunch] automatize the use of Docker. The former is perfectly compatible with the described workflow and we recommend it to users  who are not comfortable with command-line use of Docker. `holepunch` uses [binder](https://mybinder.org) [@binder2018] to move the analysis to the cloud, so that no local installation of Docker is required. `holepunch` is well suited for simple analysis with low computational demands, because binder's  memory and computing time is limited. There are several alternatives to Docker that manage dependencies on R packages. [`packrat`](https://rstudio.github.io/packrat/) [@packrat] is a way to freeze package version via local copies of packages in the project, but it does not gurantee a given base R version or system dependencies. Similar approaches are taken by [`jetpack`](https://github.com/ankane/jetpack) [@jetpack], [`miniCRAN`](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint]. The package [`reprex`](https://github.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex] is also worth noting, but its scope is limited.

>>>>

Git also makes it particularly easy to share and collaborate on a project with other researchers. A popular service for sharing materials via Git is [GitHub](https://github.com). Alternatively, institutions can host an equally feature-rich open-source software called [GitLab](https://gitlab.com) themself, avoiding the reliance on commercial service providers. Just sharing Git repositories on GitHub with the public is always free, private repositories (only visible to persons you invite) are [free for researchers](https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount). After creating a user account, one can create a new repository and GitHub provides information on how to upload your repository from the terminal, e.g., for our repository (here with user name 'aaronpeikert' and repository name 'reproducible-research'):

@@

While our approach was designed to scale well with the complexity of a computing-intense project, we realise that this flexibility may not be straightforward to integrate into a researcher's everyday workflow. There are R packages that implement parts of our workflow and, thus, lower the threshold for adoption when the full flexibility provided by our workflow is not needed. The use of R Markdown within a project, tracked with Git can be simplified with the [`workflowR`](https://github.com/jdblischak/workflowr) package [@workflowr]. The [`drake`](https://github.com/ropensci/drake) package [@drake] is directly inspired by Make but can only handle dependencies within R. The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [`holepunch`](https://karthik.github.io/holepunch/) package [@holepunch] automatize the use of Docker. The former is perfectly compatible with the described workflow and we recommend it to users  who are not comfortable with command-line use of Docker. `holepunch` uses [binder](https://mybinder.org) [@binder2018] to move the analysis to the cloud, so that no local installation of Docker is required. `holepunch` is well suited for simple analysis with low computational demands, because binder's  memory and computing time is limited. There are several alternatives to Docker that manage dependencies on R packages. [`packrat`](https://rstudio.github.io/packrat/) [@packrat] is a way to freeze package version via local copies of packages in the project, but it does not gurantee a given base R version or system dependencies. Similar approaches are taken by [`jetpack`](https://github.com/ankane/jetpack) [@jetpack], [`miniCRAN`](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint]. The package [`reprex`](https://github.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex] is also worth noting, but its scope is limited.
